---
title: "Get Started"
vignette: >
  %\VignetteIndexEntry{Build OmniPathR databases}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
author: "`r 'Author: Matthew Hung'`"
date: "`r paste('Last Updated:', format(Sys.Date(), '%Y/%m/%d'))`"
output:
  html_document:
    code_folding: hide
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```


Now let's load the package:

```{r}
library(plotr)
library(ggplot2)
```

# Theme Functions

The plotr package includes several theme functions to quickly style your ggplot2 visualizations:

## theme_text()

`theme_text()` provides consistent text styling with Helvetica font and appropriate sizing:

```{r}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  theme_text() +
  labs(title = "theme_text() Example")
```

## theme_border()

`theme_border()` provides a clean theme with borders:

```{r}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  theme_border() +
  theme_text() +
  labs(title = "theme_border() Example")
```

## theme_line()

`theme_line()` provides a minimal theme with axis lines instead of full borders:

```{r}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  theme_line() +
  theme_text() +
  labs(title = "theme_line() Example")
```

## Additional Themes

Other themes include:

- `theme_noaxisticks()`: Theme without axis ticks
- `theme_gridlines()`: Lighter gridlines styling
- `facet_aes()`: Enhanced facet_wrap aesthetics
- `umap_aes()`: Special aesthetics for dimension reduction plots

# Color Palettes

The plotr package provides access to various color palettes:

## Available Palettes

You can use the `display_palettes()` function to see all available palettes:

```{r fig.height=8}
display_palettes()
```

## Using Palettes

You can use the `get_palette()` function to retrieve colors from a palette:

```{r}
# Get colors from a package palette
kelly_colors <- get_palette(pal_kelly_20, n = 5)
print(kelly_colors)

# Use with ggplot
ggplot(mtcars, aes(wt, mpg, color = factor(cyl))) +
  geom_point(size = 3) +
  scale_color_manual(values = kelly_colors) +
  theme_line() +
  theme_text() +
  labs(title = "Kelly Colors Palette Example")
```

The `get_palette()` function also works with viridis and RColorBrewer palettes:

```{r}
# Using viridis
ggplot(mtcars, aes(wt, mpg, color = factor(cyl))) +
  geom_point(size = 3) +
  scale_color_manual(values = get_palette("viridis", n = 3)) +
  theme_line() +
  theme_text() +
  labs(title = "Viridis Palette Example")

# Using RColorBrewer
ggplot(mtcars, aes(wt, mpg, color = factor(cyl))) +
  geom_point(size = 3) +
  scale_color_manual(values = get_palette("Blues", n = 3)) +
  theme_line() +
  theme_text() +
  labs(title = "Blues Palette Example")
```

# Plot Functions

## plot_percent()

The `plot_percent()` function creates percentage/frequency bar plots:

```{r}
# Convert variables to factors first
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$am <- as.factor(mtcars$am)

# Basic percentage plot
plot_percent(mtcars, group.by = "am", variable = "cyl")
```

### Customized plot_percent

The function has many customization options:

```{r}
# With custom colors
plot_percent(
  mtcars, 
  group.by = "am", 
  variable = "cyl",
  cols = get_palette("viridis", n = 3),
  bar.width = 0.7,
  bar.color = "black",
  y.label = "Percentage"
)
```

### Faceted Percentage Plot

You can also create faceted plots:

```{r fig.width=10}
# Convert vs to factor
mtcars$vs <- as.factor(mtcars$vs)

# Faceted plot
plot_percent(
  mtcars, 
  group.by = "am", 
  variable = "cyl",
  facet.by = "vs",
  cols = get_palette(pal_kelly_20, n = 3)
)
```

### Horizontal Percentage Plot

You can change the orientation to horizontal:

```{r}
plot_percent(
  mtcars, 
  group.by = "am", 
  variable = "cyl",
  orientation = "horizontal",
  theme = "border",
  cols = get_palette(pal_zissou_5, n = 3),
  x.label = "Transmission Type"
)
```

# Specialized Visualizations

## UMAP Aesthetics

For dimension reduction plots, you can use the `umap_aes()` function:

```{r fig.height=6, fig.width=6}
# Create sample UMAP data
set.seed(123)
umap_data <- data.frame(
  UMAP1 = rnorm(100),
  UMAP2 = rnorm(100),
  cluster = sample(LETTERS[1:5], 100, replace = TRUE)
)

# Plot with umap_aes
ggplot(umap_data, aes(UMAP1, UMAP2, color = cluster)) +
  geom_point(size = 3) +
  theme_text() +
  scale_color_manual(values = get_palette(pal_brewerplus_41, n = 5)) +
  labs(title = "UMAP Example") +
  umap_aes()
```
